---
title: "p8105_hw1_rl3616"
author: "Ruipeng Li"
date: "`r format(Sys.Date())`"
output: github_document
---
```{r setup, message=FALSE}
library(moderndive)               #Attach the moderndive package
library(ggplot2)                  #Load in order to use ggplot
data("early_january_weather")     #Load a built-in dataset
```

# Problem 1
This dataset has "`r nrow(early_january_weather)`" rows and "`r ncol(early_january_weather)`" columns.

Variables include: "`r paste(names(early_january_weather), collapse = ", ")`".

Mean temperature (F): "`r round(mean(early_january_weather$temp, na.rm = TRUE), 2)`".

```{r}
p <- ggplot(early_january_weather, 
  aes(x = time_hour, y = temp, color = humid)) +      #initialize and setting up a ggplot object
  geom_point(alpha = 0.7) +                           #setting point transparency
  scale_x_datetime(
    date_breaks = "2 day",
    date_labels = "%d"
  ) +                                                 #customize x-axis scale for datetime data
  labs(
    title = "Temperature vs Time (colored by Humidity)",
    x = "Time in January (hourly data)",
    y = "Temperature (F)",
    color = "Humidity"
  ) +                                                 #set plot labels
  theme_minimal()                                     #apply a theme I love♡

p                                                     #show the garph

```

In the chart, each vertical line represents 0:00 a.m. of that day (e.g., the line labeled "02" corresponds to 0:00 a.m. on January 2). We can see that temperature peaks typically occur during the daytime between two midnights, while temperature troughs usually appear near 0:00 a.m. each day. 

Lighter blue dots indicate higher humidity.High nighttime humidity often leads to larger diurnal temperature swings (e.g., from January 6 to 11) and can cause sharp temperature drops (e.g., around January 15). Meanwhile, persistently high humidity tends to reduce diurnal temperature variation (e.g., from noon on January 11 to January 14).

```{r}
ggsave("temp_vs_time_humid.png", plot = p)
```
Here's the saving progress.

